An asymptotically normal out-of-sample test of equal predictive accuracy for nested models

### 2013-02-14 Paper.tex
Finish edits to bootstrap normality
Makes additional changes to the statement of Theorem 3 and to the proof of
asymptotic normality for the bootstrap process, so that I can invoke de Jong
and Davidson's proof for the convergence of the second moment.

\begin{thm}\label{res:3}
  Suppose the following conditions hold:
  \begin{enumerate}
  \item Let $\{y_t, z_t\}$ be an $L_2$-\textsl{\ned} process of size
    $-\frac12$ on the series $V_t$; $V_t$ is strong mixing of size
    $-\frac{r}{r-2}$ or uniform mixing of size $-\frac{r}{2r-2}$ with
    $r > 2$.\footnote{For uniform mixing processes, $r = 2$ is also
      allowed as long as $\sup_{\beta \in N} \psi_t(\beta)^2$ and
      $\sup_{\beta \in N} (\partial \psi_t(\beta) / \partial \beta)^2$
      are uniformly integrable; these terms are defined later in these
      assumptions.}  Moreover, the \ned-coefficients are uniformly
    bounded.
  \item The sequence $(y_1^{*}, z_t^*),\dots,(y_T^{*}, z_T^*)$ is
    generated by the moving blocks, circular blocks, or stationary
    bootstrap with block lengths drawn from the geometric
    distribution.  The (expected) block length $b$ satisfies $b \to
    \infty$ and $\frac{b}{P} \to 0$.
  \item The estimator $\hat{\beta}_t$ of $\beta_0$ is estimated with a
    recursive, rolling, or fixed estimation window and satisfies
    $\hat{\beta}_{t} - \beta_{0} = \hat{B}_{t} H_t$; $\hat{B}_{t}$ is
    a sequence of $k$ by $q$ matrices such that $\sup_t |\hat{B}_t -
    B| \to^p 0$, $\sup_t |\hat{B}_t^{*} - B^{*}| \to^p 0$, and $B^{*}
    = B + o_p$; $H_{t}$ is a sequence of $q$-vectors such that 
    \begin{equation}
      H_{t} = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s} & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s} & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s} & \text{fixed window} \\
      \end{cases}
    \end{equation}
    $h_{s}(\beta) = h(y_{s}, z_{s}, \beta)$, $h_s = h_s(\beta_0)$, and
    $\E h_{s} = 0$ for all $s$; and $\frac12 \sum_{t=1}^T h_t^* = 0$
    a.s. where $h_t^* = h(y_s^*, z_s^*, \hat\beta_T)$.  Also define
    \begin{equation}
      H_{t}^* = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s}^* & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s}^* & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s}^* & \text{fixed window} \\
      \end{cases}
    \end{equation}
  \item Let $f_{t}(\beta) = f(y_{t}, z_{t}, \beta)$ be an $l$-vector;
    the asymptotic variance matrix of $\bar{f}(\beta_0)$ is uniformly
    positive definite.
  \item Define $\psi_t(\beta) = (f_t(\beta), h_t(\beta))$;
    $\psi_t(\beta)$ is covariance stationary and is continuously
    differentiable in an open neighborhood $N$ of $\beta_0$.  Also,
    $\E( \sup_{\beta \in N} |\psi_t(\beta)|)^r$ and $\E\sup_{\beta \in
      N} |\frac{\partial \psi_t(\beta)}{\partial \beta}|^2$ are
    uniformly bounded.
  \item $R, P \to \infty$ as $T \to \infty$.
  \end{enumerate}
  Then
  \begin{equation}
    \pr[\sup_x | \pr^*[\sqrt{P} (\bar{f}^* - \bar{f}(\hat{\beta}_{T}))
        \leq x] - \pr[\sqrt{P}( \bar{f} - \E \bar{f}(\beta_0)) \leq x] | >
      \epsilon] \to 0
  \end{equation}
  for all $\epsilon > 0$.
\end{thm}

@@

\begin{proof}[Proof of Lemma~\ref{res:a3}]
  Note that $\psi_t$ satisfies a \clt, \citep[Theorem~2]{Jon:97}; the
  bootstrap average obeys a \clt\ as well from
  \citet[Theorem~2]{Cal:11d}, so it suffices to prove that
  \[ \tfrac{ 1 }{ P } \var^* \Big( \sum_{t=R+1}^T \psi_t^*( \hat
  \beta_T ) \Big) - \tfrac{ 1 }{ P } \var \Big( \sum_{t=R+1}^T \psi_t(
  \beta_0 ) \Big) \to^p 0. \] I will prove this result for the Moving
  Blocks Bootstrap---the proof for other block bootstraps is similar
  and follows as in \cite{Cal:11d}.

  \newcommand{\su}{\tfrac{ 1 }{ T - b } \sum_{ \tau = 0 }^{ T - b - 1
    } \tfrac{ 1 }{ b }\sum_{ s, t = \tau + 1 }^{ \tau + b }}

  Note that
  \begin{equation*}
  \tfrac{ 1 }{ P } \var^* \Big( \sum_{ t = R + 1 }^T \psi_t^*( \hat \beta_T ) \Big)
  = \su \hat \psi_s( \hat \beta_T ) \hat \psi_t( \hat \beta_T )'
     - \tilde \psi( \hat \beta_T ) \tilde \psi( \hat \beta_T )'
  \end{equation*}
  where
  \begin{equation*}
    \tilde \psi( \hat \beta_T )
    = \tfrac{ 1 }{ T - b } \sum_{ t = b + 1 }^{ T - b } \psi_t( \hat \beta_T )
       + \tfrac{ 1 }{ T - b } \sum_{ t = 1 }^b \tfrac{ t }{ b }
         \big( \psi_t( \hat \beta_T ) + \psi_{ T - t + 1 }( \hat \beta_T ) \big)
  \end{equation*}
  Since $\bar \psi( \hat \beta_T ) = 0$ by construction and (finish explanation), we have
  \begin{align*}
    \tilde \psi( \hat \beta_T )
    &= \tfrac{ b }{ T } \tfrac{ 1 }{ T - b } \sum_{ t = b + 1 }^{ T - b } \psi_t( \hat \beta_T )
       + \tfrac{ 1 }{ T - b } \sum_{ t = 1 }^b 
       \Big( \tfrac{ t - b }{ b } - \tfrac{ b }{ T } \Big)
       \big( \psi_t( \hat \beta_T ) + \psi_{ T - t + 1 }( \hat \beta_T ) \big) \\
    &= o_p(1).
  \end{align*}
  Moreover,
  \begin{equation*}
    \su \hat\psi_s \hat\psi_t' \to^p \var \Big( \sum_{t=R+1}^T \psi_t \Big)
  \end{equation*}
  by \citet[Theorem 2.2]{JoD:00}.  To see this, observe that 
  \begin{equation*}
    \su \hat\psi_s \hat\psi_t' 
    = \tfrac{1}{T} \sum_{s,t=1}^T \hat\psi_s \hat\psi_t' \;
      \big(1 - \tfrac{\lvert s - t \rvert}{b}\big)^+ + term.
  \end{equation*}
  The first term satisfies most of \citepos{JoD:00} Assumptions by
  construction and it suffices to prove their condition (2.9) holds,
  namely that
  \begin{equation}\label{eq:2}
    \sup_{\beta \in N} \Bigg\lVert \tfrac{1}{T} \sum_{t=1}^T e^{i \xi t/b} 
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} - 
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg) 
    \Bigg\rVert_2 \to 0
  \end{equation}
  for all $i$, $j$, and $\xi$.  I will show that \eqref{eq:2} holds
  for the real part of the expression; the proof for the complex part
  is similar.

  For any fixed $\beta \in N$,
  \begin{equation*}
    \tfrac{1}{T} \sum_{t=1}^T \cos( \xi t / b )
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} - 
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg) 
  \end{equation*}
  obeys an \lln\ (give reference and explanation -- you need to show
  that the derivatives are \ned\ as well.) and convergence in $L_2$
  follows from uniform integrability.

  TODO: add stochastic equicontinuity conditions.
\end{proof}

>>>>

\begin{thm}\label{res:3}
  Suppose the following conditions hold:
  \begin{enumerate}
  \item The estimator $\hat{\beta}_t$ of $\beta_0$ is estimated with a
    recursive, rolling, or fixed estimation window and satisfies
    $\hat{\beta}_{t} - \beta_{0} = \hat{B}_{t} H_t$; $\hat{B}_{t}$ is
    a sequence of $k \times q$ matrices such that $\sup_t |\hat{B}_t -
    B| \to^p 0$; $H_{t}$ is a sequence of $q$-vectors such that
    \begin{equation}
      H_{t} = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s}(\beta_0) & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s}(\beta_0) & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s}(\beta_0) & \text{fixed window} \\
      \end{cases}
    \end{equation}
    and $\E h_{s}(\beta_0) = 0$ for all $s$.
  \item Define $\psi_t(\beta) = (f_t(\beta), h_t(\beta))$;
    $\psi_t(\beta)$ is covariance stationary and twice continuously
    differentiable with bounded second derivative in an open
    neighborhood $N$ of $\beta_0$.  Also, $\E \sup_{\beta \in N}
    |\psi_t(\beta)|^r$ and $\E\sup_{\beta \in N} |\frac{\partial
      \psi_t(\beta)}{\partial \beta'}|^r$ are uniformly bounded ($r$ is
    defined below).
  \item Both $\psi_t(\beta)$ and $\frac{\partial}{\partial \beta}
    \psi_t(\beta)$ are $L_2$-\textsl{\ned} of size $-\frac12$ on the
    series $V_t$ with bounded \ned-coefficients for $\beta$ uniformly
    in $N$; $V_t$ is strong mixing of size $-\frac{r}{r-2}$ or uniform
    mixing of size $-\frac{r}{2r-2}$ with $r > 2$.\footnote{For
      uniform mixing processes, $r = 2$ is also allowed as long as
      $\sup_{\beta \in N} \psi_t(\beta)^2$ and $\sup_{\beta \in N}
      (\partial \psi_t(\beta) / \partial \beta)^2$ are uniformly
      integrable.}
  \item The sequence $\psi_1^{*},\dots,\psi_T^{*}$ is generated by the
    moving blocks, circular blocks, or stationary bootstrap with block
    lengths drawn from the geometric distribution.  The (expected)
    block length $b$ satisfies $b \to \infty$ and $\frac{b}{P} \to 0$.
  \item The bootstrapped estimator satisfies $\sup_t |\hat{B}_t^{*} -
    B^{*}| \to^p 0$, $B^{*} = B + o_p$ and $\frac1T \sum_{t=1}^T h_t^* = 0$ a.s.
  \item The asymptotic variance matrix of $\bar{f}(\beta_0)$ is
    uniformly positive definite.
  \item $R, P \to \infty$ as $T \to \infty$.
  \end{enumerate}
  Then
  \begin{equation}
    \pr[\sup_x | \pr^*[\sqrt{P} (\bar{f}^* - \bar{f}(\hat{\beta}_{T}))
        \leq x] - \pr[\sqrt{P}( \bar{f} - \E \bar{f}(\beta_0)) \leq x] | >
      \epsilon] \to 0
  \end{equation}
  for all $\epsilon > 0$.
\end{thm}

@@

\begin{proof}[Proof of Lemma~\ref{res:a3}]
  Note that both $\psi_t(\beta_0)$ and $\psi_t^*(\hat\beta_T)$ satisfy
  \clt s, (\citealp[Theorem~2]{Jon:97} and
  \citealp[Theorem~1]{Cal:11d}) so it suffices to prove that
  \[ \tfrac{ 1 }{ P } \var^* \Big( \sum_{t=R+1}^T \psi_t^*( \hat
  \beta_T ) \Big) - \tfrac{ 1 }{ P } \var \Big( \sum_{t=R+1}^T \psi_t(
  \beta_0 ) \Big) \to^p 0. \]
  I will prove this result for the Circular Block Bootstrap
  \citep{PoR:92}---the proof for other block bootstraps is similar and
  follows as in \cite{Cal:11d}.

  \newcommand{\su}{\tfrac{ 1 }{ b T } \sum_{ \tau = 1 }^{ T }%
  \sum_{ s, t \in I( \tau ) }}

  Since $\E^* \psi_t^*( \hat \beta_T ) = 0$ by construction, we have
  \begin{equation}\label{eq:3}
  \tfrac{ 1 }{ P } \var^* \Big( \sum_{ t = R + 1 }^T \psi_t^*( \hat \beta_T ) \Big)
  = \su \hat \psi_s( \hat \beta_T ) \hat \psi_t( \hat \beta_T )'
  \end{equation}
  almost surely, with
  \begin{equation*}
    I(\tau) =
    \begin{cases}
      \{ \tau, \dots, \tau + b - 1 \} & \text{if } \tau \leq T - b + 1 \\
      \{ \tau, \dots, T \} \cup \{ 1, \dots, \tau - ( T - b + 1 ) \}
      & \text{if } \tau > T - b + 1.
    \end{cases}
  \end{equation*}
  We can rewrite the right side of \eqref{eq:3} as
  \begin{multline}\label{eq:6}
    \su \psi_s(\hat\beta_T) \psi_t(\hat\beta_T)'
    = \tfrac{1}{T} \sum_{s,t=1}^T \psi_s(\hat\beta_T) \psi_t(\hat\beta_T)' \;
      \big(1 - \tfrac{\lvert s - t \rvert}{b}\big)^+ \\
      + \tfrac{1}{bT} \sum_{s = 1}^{b - 1} \sum_{t = T - b + s}^{T}
        (\psi_s(\hat\beta_T) \psi_t(\hat\beta_T)' +
         \psi_t(\hat\beta_T) \psi_s(\hat\beta_T)') \;
        \big(1 - \tfrac{\lvert s - t + T\rvert}{b}\big)^+.
  \end{multline}

  The first term of \eqref{eq:6} can be shown to converge in
  probability to $(1/P) \var \sum_{t=R+1}^T \psi_t( \beta_0 )$
  by Theorem 2.2 of \citet{JoD:00}.  Most of the necessary conditions
  hold by assumption and it suffices to prove their condition (2.9)
  holds, namely that
  \begin{equation}\label{eq:2}
    \sup_{\beta \in N} \Bigg\lVert \tfrac{1}{T} \sum_{t=1}^T e^{i \xi t/b} 
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} -
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg) 
    \Bigg\rVert_2 \to 0
  \end{equation}
  for all $i$, $j$, and $\xi$.  Now, for any fixed $\beta \in N$,
  \begin{equation*}
    \tfrac{1}{T} \sum_{t=1}^T \cos( \xi t / b )
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} - 
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg)
    \to^p 0
  \end{equation*}
  and
  \begin{equation*}
    \tfrac{1}{T} \sum_{t=1}^T \sin( \xi t / b )
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} - 
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg)
    \to^p 0
  \end{equation*}
  by a mixingale \lln\ \citep{Dav:93}.\footnote{Remember that \ned\
    processes are also mixingales.  See, for example, \citet[Section
    17.2]{Dav:94}.}  Convergence in $L_2$ for each $\beta \in N$ then
  follows from the existence of the $r$th moment and uniform
  convergence is a consequence of the bounded second derivative of
  $\psi_t$.

  The conclusion then holds if the second term of~\eqref{eq:6} is
  $o_p(1)$.  We have, for each element of that matrix,
  \begin{multline}
    \Big| \tfrac{1}{bT} \sum_{s = 1}^{b - 1} \sum_{t = T - b + s}^{T}
        \psi_{is}(\hat\beta_T) \psi_{jt}(\hat\beta_T)
        \big(1 - \tfrac{\lvert s - t + T\rvert}{b}\big)^+ \Big| \leq\\
     \tfrac{1}{T} \sum_{s = 1}^{b - 1} |\psi_{is}(\hat\beta_T)|
         \times\tfrac{1}{b} \sum_{t = T - b + 1}^{T} | \psi_{jt}(\hat\beta_T) |
  \end{multline}
  almost surely.  Now, for any positive $\delta$,
  \begin{equation*}
    \pr \Bigg[ \tfrac{1}{T} \sum_{s = 1}^{b - 1} | \psi_{is}( \hat \beta_T ) |
              > \delta \Bigg] \leq 
    \pr \Bigg[\tfrac{1}{T} \sum_{s = 1}^{b - 1} \sup_{ \beta \in N } | \psi_{is}( \beta ) |
              > \delta \Bigg] +
    \pr \big[ \hat \beta_T \notin N \big].
  \end{equation*}
  Both terms converge to zero by assumption, so $\tfrac{1}{T} \sum_{s
    = 1}^{b - 1} |\psi_{is}(\hat\beta_T)| = o_p(1)$.  A similar
  argument shows that $\tfrac{1}{b} \sum_{t = T - b + 1}^{T} |
  \psi_{jt}(\hat\beta_T) | = O_p(1)$, completing the proof.
\end{proof}
